A Review of Neutron Imaging

### 2020-10-11 neutron-imaging/neutron.tex
WIP 11Oct2020

\section{Propagation of neutron beams}\label{s7}
Typical neutron sources have size of the order of $10$ cm in all directions.
They are enclosed in moderators to produce a quasi-Maxwell distribution
\cite{anderson2008neutron}. Their brightness is defined as the flux of neutrons
per solid angle per unit wavelength. The flux itself is the number of neutrons
crossing a unit area. One can treat the neutrons using the principles of
geometrical optics. The phase space of analytical mechanics can be carried over
to optics and is called etendue. Consider an infinitesimal surface element 
$dS$ with normal $\hat{n}$ pointing in a medium of refractive index $n$. If
the surface $dS$ is crossed by neutrons confined to a solid angle $d\Omega$
then the etendue of neutron beam is
\begin{equation}\label{s7e1}
dG = n^2\hat{n}\cdot\hat{n}_b dSd\Omega,
\end{equation}
where $\hat{n_b}$ is the normal along the axis of the cone of $d\Omega$ 
\cite{markvart2007thermodynamics}. The way the volume of the phase space is
conserved, one can also conclude that the volume of an etendue is conserved. 
The optical analogue of Liouville's theorem places a limit on the efficiency
of a neutron source \cite{anderson2008neutron}, 
\cite{markvart2007thermodynamics}. If one want to focus a neutron beam from
a source of size $w_1$ on a spot of size $w_2 < w_1$ then one must allow the
beam to diverge in directions normal to the line joining the source and the
spot.

@@

\section{Neutron diffraction by crystals}\label{s8}
\section{2-d neutron imaging}\label{s9}
\section{Neutron tomography}\label{s10}
\section{Imaging of biologically important structures}\label{s11}
\section{Larmor precession based imaging}\label{s12}
\section{Polarization based imaging}\label{s13}
\section{Conclusion}\label{s14}
\bibliography{neutron}
\bibliographystyle{plain}
\end{document}

>>>>

\section{Optics of neutron beams}\label{s7}
Typical neutron sources have size of the order of $10$ cm in all directions.
They are enclosed in moderators to produce a quasi-Maxwell distribution
\cite{anderson2008neutron}. Their brightness is defined as the flux of neutrons
per solid angle per unit wavelength. The flux itself is the number of neutrons
crossing a unit area. One can treat the neutrons using the principles of
geometrical optics. The phase space of analytical mechanics can be carried over
to optics and is called etendue. Consider an infinitesimal surface element 
$dS$ with normal $\hat{n}$ pointing in a medium of refractive index $n$. If
the surface $dS$ is crossed by neutrons confined to a solid angle $d\Omega$
then the etendue of neutron beam is
\begin{equation}\label{s7e1}
dG = n^2\hat{n}\cdot\hat{n}_b dSd\Omega,
\end{equation}
where $\hat{n_b}$ is the normal along the axis of the cone of $d\Omega$ 
\cite{markvart2007thermodynamics}. The way the volume of the phase space is
conserved, one can also conclude that the volume of an etendue is conserved. 
The optical analogue of Liouville's theorem places a limit on the efficiency
of a neutron source \cite{anderson2008neutron}, 
\cite{markvart2007thermodynamics}. If one want to focus a neutron beam from
a source of size $w_1$ on a spot of size $w_2 < w_1$ then one must allow the
beam to diverge in directions normal to the line joining the source and the
spot.

@@

The existence of spin states of neutrons allows us to describe beams of 
neutrons by their polarization states. Being a spin-half particle, a neutron
can exist in one of $|+\rangle$ of $|-\rangle$ states. If a beam has $N_+$
neutrons in $|+\rangle$ and $N_-$ neutrons in $|-\rangle$ states then its
polarization is defined to be
\begin{equation}\label{s7e5}
P = \frac{N_+ - N-}{N_+ + N_-}.
\end{equation}
It is equally common to express beam polarization in terms of flipping ratio
as
\begin{equation}\label{s7e6}
F = \frac{N_+}{N_-} = \frac{1 - P}{1 + P}.
\end{equation}
When the nuclear and the magnetic structure factors of a crystal are equal in
magnitude, the total structure factor vanishes for one of the two spin states.
As a result, the Bragg reflection from the crystal plance produces a beam with
almost all neutrons in the remaining polarization states. The most commonly
used crystal to achieve neutron polarization is the [111] reflection of 
Heusler alloy \cite{anderson2008neutron}. This technique is most suitable for
steady state source of neutrons \cite{williams1988polarized}. In the case of
a pulsed beam of neutrons it is common to use resonance absorption filters.
When an unpolarized neutron beam whose energy corresponds to a nuclear
resonance is shone on a target of polarized nuclear spins, one of the 
neutron spins is absorbed while the other one is transmitted 
\cite{williams1988polarized}.

In order to use a neutron beam for imaging the neutrons emerging from a source
must be collimated into a stream of particles traveling mostly along a single
direction. Being electrically neutral, they cannot be focused using an electric
field. Their small magnetic moment requires a very large field gradient to 
change their direction of propagation. A simpler way of achieving collimation
is to collect neutrons in a tube lined with a material which is opaque to 
neutrons. Collimators have a small entrance aperture and a larger exit allowing
a large neutron flux.

\section{2-d neutron imaging}\label{s8}
2-d neutron imaging is similar to the more commonly known 2-d X-ray imaging.
When X-rays are incident on a target and when then the intensity of the 
transmitted beam is analyzed one observes that different portions of the
latter have different intensities. This is because different portions of the
transmitted beam have emerged from different parts of the target each one
having their own attenuation properties. 2-d neutron imaging works in exactly
the same manner. Attenuation of X-rays depends on the local electronic
properties of the target while that of neutron depends on the local nuclear
properties. That is why, the two imaging techniques are often considered to
provide complementary information about the target. X-rays provide information
about the bonding or structural properties while neutrons indicate the 
elemental composition. The set up for 2-d neutron imaging is quite simple. It 
has a source of neutrons surrounded by a moderator to thermalize them, a 
collimator to coax the particles along a single direction and a detector to
visualize the image.

The intensity of the incident and the transmitted beams is given by equation
\eqref{s4e2}. If $I_0$ and $I$ are the intensities of the incident and the
transmitted beams then the degree of darkness of the film used to visualize
the image is
\begin{equation}\label{s8e1}
D_e = \ln\frac{I_0}{I}.
\end{equation}
The film brightens most where the transmitted intensity is the highest. In the
case of a digital receptor, the grey level value $G$ is
\begin{equation}\label{s8e2}
G = Cn + G_0,
\end{equation}
where $C$ is a constant, $n$ is the neutron density and $G_0$ is another 
constant that indicates the gray level due to noise in the circuits. The 
presence of $G_0$ suggests that $G \ne 0$ even in the absence of neutron flux.
This method suffers from the disadvantage of catching signals not only from
the transmitted neutrons but from other ambient radiation as well. As a result,
the images often are blurred. Nevertheless, it is used in situation where the
sharpness of the image is not critical. For example, in the non-destructive
testing in materials, it sufficient to know if the material has contaminants,
defects, lubrication etc. It is also used in the investigation of two-phase
flows, flows through porous media \cite{anderson2008neutron}, detection of
hydrgenous materials and moisture \cite{richards2004neutron}.

\section{Neutron tomography}\label{s9}
Tomography is an imaging technique in which the image of a target is constructed
by combining the images of its slices. The most usage is in medical imaging
in the form of CT (Computerized Tomography) scans. X-rays are used over a 
large number of slices of the human body. The `image' produced by each slice
is combined with others to get complete information of the portion of interest.
The way 2-d neutron imaging is an adaptation of 2-d X-ray imaging, neutron
tomography is applying the principles of X-ray tomography to neutron imaging. 
Consider a target $T$ whose image we are are interested in. We place the origin
of the cartesian coordinate system at a suitable point in $T$. We consider $T$
to be composed of a large number of slices whose normal is along the $z$-axis.
We know from equation \eqref{s4e2} that the attenuation of intensity of an
incident neutron beams by a slice $S$ is described by the equation
\begin{equation}\label{s9e1}
I = I_0\exp\left(-\int_S\mu(x, y)ds\right).
\end{equation}
Since the target is not a homogenous object, we cannot consider it to have a
uniform attenuation $\mu$ as we did in equation \eqref{s4e2}. The parameter
$s$ is the coordinate along the direction of propagation of neutrons in the
target. We can rearrange equation \eqref{s9e1} and define a function
\begin{equation}\label{s9e2}
p(t, \theta) = \ln\left(\frac{I}{I_0}\right) = -\int_S\mu(x, y)ds.
\end{equation}
The coordinate $t$ runs perpendicular to the coordinate $s$. One can write 
this in terms of an integral that extends over all $s$ as
\begin{equation}\label{s9e3}
p(t, \theta) = 
-\int_{-\infty}^\infty \mu(x, y)\delta(t - x\cos\theta - y\sin\theta)ds.
\end{equation}
The function $p(t, \theta)$ is called the Radon transform of the slice. It is
taken along the line $t = x\cos\theta + y\sin\theta$, expressed in the 
normal form \cite{gonzales1992digital}. It is the projection of the function
$f$ on the line $t = x\cos\theta + y\sin\theta$. It can be shown using the
Fourier Slice Theorem \cite{bracewell1956strip} that a Fourier transform of
$p(t, \theta)$ is the same as the slice of the 2-d Fourier transform of the
target. Thus, taking the 2-d inverse Fourier transform of the result of the
first operation will give us the original function. The number of such
projections is determined by Shannon's sampling theorem.

Neutron tomography has the same advantages over neutron radiography that CT
scans has over X-ray images. Neutron radiography gives information about the
total attenuation of a neutron beam along its path through the target. It 
cannot give information about the actual distribution of materials across that
path. One of the earliest use cases of neutron tomography was in finding 
concentration of hydrogen in metal castings. The presence of hydrogen is an
indicator of the loss of mechanical properties due to embrittlement. We know
that the collision cross-section of neutron with hydrogen is very large and
that neutron imaging is a powerful tool to investigate hydrogenous materials. 
Neutron tomography techniques were able to give accurate information 
about hydrogen concentration as low as $200$ ppm \cite{richards2004neutron}.
Tomographic techniques are also suitable for investigating objects that do not
have a substantial plane surface like the blade of a turbine or an aircraft 
engine \cite{richards2004neutron}.

\section{Imaging of biologically important structures}\label{s10}
\section{Larmor precession based imaging}\label{s11}
\section{Polarization based imaging}\label{s12}
\section{Conclusion}\label{s13}
\bibliography{neutron}
\bibliographystyle{plain}
\end{document}
